---
title: "Reclustering of most immature HSPCs"
output: html_document
date: "2024-05-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load libraries
library(tidyverse)
library(Seurat)
library(condiments)
library(slingshot)
library(destiny)
library(dittoSeq)
library(dplyr)
library(tidyr)
library(missMDA)
library(dismo)
library(RColorBrewer)
library(gridExtra)
library(tradeSeq)
```

```{r}
#load input data
#Input data: Cells from cluster 0 1 5 6 7 9 17 18 from previous analysis
dat <- readRDS ("trajectory_input_cells.rds")

# specify the default assays as "mrna_cca".
DefaultAssay(dat)<-"mrna_cca"
```

The below samples are the sorted immature populaiton (spike-in). They are removed from the downstream analysis.

old18_CD34CD38_hs
old12_CD34CD38_hs
old4_CD34CD38_hs
old19_CD34CD38_hs
midage14_hs
midage2_hs
midage11_hs
midage8_hs
midage10_hs
midage15_hs

After removing the spike-in samples, PCA and clustering are recalculated.


## 1. Remove spike-in
```{r}
# change identity
Idents(dat)<-dat$sample

# subset samples
dat <- subset(dat, idents = c("old18_CD34CD38_hs","old12_CD34CD38_hs","old4_CD34CD38_hs","old19_CD34CD38_hs", "midage14_hs","midage2_hs","midage11_hs","midage8_hs","midage10_hs","midage15_hs"), invert=T)

# # double check the subset
# dat$sample %>% unique()
```

## 2. Recalculate PCA
```{r}
# Recalculate PCA and find optimal number of PCs
# the recalculated PCA is stored in the default - dat@reductions$pca
dat <- RunPCA(dat, npcs = 50, verbose = FALSE, approx = TRUE)

# get features with variance > 0
dat_varFeature <- PrepDR(object = dat, genes.use = VariableFeatures(object = dat), use.imputed = F, assay.type = "mrna_cca")

# # find the optimal number of PCs

nPC_dat <- PCA_estimate_nPC(dat_varFeature, whereto="rdata/01_nPC.rds", by.nPC=1, to.nPC = 50)
```

## 3. Recluster cells with resolution 0.3 / 0.4 / 0.6
```{r}
# find neighbours
dat <- FindNeighbors(dat, dims = 1:nPC_dat, verbose = FALSE)

# find clusters under resolution 0.3/0.4/0.6
# graph used: mrna_cca_snn
dat <- FindClusters(dat, resolution = c(0.3, 0.4, 0.6), verbose = FALSE)

# UMAP 
dat <- RunUMAP(dat, dims = 1:nPC_dat)

# Plot UMAP with resolution 0.3/0.4/0.6
pdf("result/01_res_0.3_0.4_0.6.pdf", width = 20)
p1 <- DimPlot(dat, group.by = "mrna_cca_snn_res.0.3")
p2 <- DimPlot(dat, group.by = "mrna_cca_snn_res.0.4")
p3 <- DimPlot(dat, group.by = "mrna_cca_snn_res.0.6")
p1|p2|p3
dev.off()

# split clusters in umap
pdf("result/02_res_0.3_0.4__0.6_split.pdf")
p1 <- DimPlot(dat, group.by = "mrna_cca_snn_res.0.3", split.by = "mrna_cca_snn_res.0.3")
p2 <- DimPlot(dat, group.by = "mrna_cca_snn_res.0.4", split.by = "mrna_cca_snn_res.0.4")
p3 <- DimPlot(dat, group.by = "mrna_cca_snn_res.0.6", split.by = "mrna_cca_snn_res.0.6")
grid.arrange(p1, p2, p3, nrow = 3)
dev.off()

saveRDS(dat,"rdata/02_dat_seuratObj.rds")
```

Resolution 0.4 is chosen for downstream analysis.

## 4.FindAllMarkers
```{r}
# switch the default assay to RNA
DefaultAssay(dat) <- "RNA"

# change identity to mrna_cca_snn_res.0.4
Idents(dat) <- dat$mrna_cca_snn_res.0.4

# perform findAllMarkers with min.pct = 0.25
findAllMarkers_res0.4_minPct0.25 <- FindAllMarkers(dat, assay = "RNA", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, test.use = "wilcox")

# perform findAllMarkers with min.pct = 0.05
findAllMarkers_res0.4_minPct0.05 <- FindAllMarkers(dat, assay = "RNA", only.pos = TRUE, min.pct = 0.05, logfc.threshold = 0.25, test.use = "wilcox")

# # save
write_csv(findAllMarkers_res0.4_minPct0.25, file = "result/03_findAllMarkers_res0.4_minPct0.25.csv")
write_csv(findAllMarkers_res0.4_minPct0.05, file = "result/03_findAllMarkers_res0.4_minPct0.05.csv")
```

Create pdf for UMAP gene expression.
```{r}
for (i in reclustered.findAllMarkers$gene) {
  require(lattice)
  pdf(paste("results/", i, "_findAllMarkers.pdf", sep = ""))
  print(FeaturePlot(recluster, features = i))
  dev.off()
 }
```

## 5. Additional ploting- dotplot with AbSeq
```{r}
 DefaultAssay(dat) <- "ab"
 
pdf("result/reclustering_yg/ab_dotplot.pdf", height = 15, width = 10)
 DotPlot(recluster, features = row.names(recluster)) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   coord_flip()
 dev.off()
```

## 6. Cluster composition - check number of cells in each cluster in each age group
```{r}
metaDat <- dat@meta.data

# number of cells in each cluster in each group
metaDat %>% 
  dplyr::select(age_group, mrna_cca_snn_res.0.4) %>% 
  group_by(age_group, mrna_cca_snn_res.0.4) %>% 
  dplyr::summarise(len = length(age_group)) %>% 
  ggplot(aes(fill=age_group, y=len, x=mrna_cca_snn_res.0.4)) + 
    geom_bar(position="dodge", stat="identity")

# stack bar plot: number of cells in each cluster in each group
metaDat %>% 
  dplyr::select(age_group, mrna_cca_snn_res.0.4) %>% 
  group_by(age_group, mrna_cca_snn_res.0.4) %>% 
  dplyr::summarise(len = length(age_group)) %>% 
  ggplot(aes(fill=age_group, y=len, x=mrna_cca_snn_res.0.4)) + 
    geom_bar(position="fill", stat="identity")
```

-- samples have varied number of cells => Normalize to total cell number of each sample
```{r}
# dot plot: Noramlized to the cell number of each sample
sample_n <- metaDat %>% 
  dplyr::select(sample) %>% 
  group_by(sample) %>% 
  dplyr::summarise(len = length(sample))

cellNo_sample_cluster <- metaDat %>% 
  dplyr::select(age_group, mrna_cca_snn_res.0.6, sample) %>% 
  group_by(age_group, mrna_cca_snn_res.0.6, sample) %>% 
  dplyr::summarise(len = length(age_group)) 

cellNo_sample_cluster %>% 
  left_join(., sample_n, by = "sample") %>% 
  mutate(percentage = len.x/len.y) %>% 
  ggplot(aes(mrna_cca_snn_res.0.6, percentage, fill = age_group)) +
  geom_dotplot(binaxis='y', stackdir='center', 
               position=position_dodge2(.6)) +
  theme_classic()

# # save
# write_csv(cellNo_sample_cluster, file = "result/04_cellNo_sample_cluster.csv")

cellNo_sample_cluster %>% 
  filter(mrna_cca_snn_res.0.4 == 0) %>% 
  .$len %>% 
  sum()
```

## 7. Save seurat object for follow-up analysis
```{r}
saveRDS(dat,"rdata/02_dat_seuratObj.rds")
```

