---
title: "01_per_cartridge_analysis"
author: "Edyta Kowalczyk"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r}
library(Seurat)
library(dplyr)
library(dittoSeq)
library(ggplot2)
```

#1. Function Generate seurat object for each cartridge with mRNA and AbSeq
```{r}
generate_seurat_object <- function(main_dir, dir_in,
                                   input_file_name,
                                   input_smk_calls) {
  #add cell indexes as row names
  expression_matrix <- read.csv(sprintf("%s/%s/%s", main_dir, dir_in,
                                        input_file_name),skip=7)
  
  expression_matrix_counts <- expression_matrix[,c(2:ncol(expression_matrix))]
  rownames(expression_matrix_counts) <- expression_matrix$Cell_Index
  
  #create abseq matrix
  expression_matrix_counts_abseq <- expression_matrix_counts[,grep("pAb",colnames(expression_matrix_counts))]
  abseq_names <- colnames(expression_matrix_counts_abseq)

  #create mrna matrix
  expression_matrix_counts_mrna <- expression_matrix_counts[,!grepl("pAb",colnames(expression_matrix_counts))]

  #adjust abseq names
  haas_abseq_names <- gsub("\\..*","-AB",abseq_names)
  colnames(expression_matrix_counts_abseq) <- haas_abseq_names

  #create seurat abseq object
  expression_matrix_counts_abseq_object <- CreateSeuratObject(t(expression_matrix_counts_abseq), project = "SeuratProject", assay = "Abseq", min.cells = 0, min.features = 0, names.field = 1, names.delim = "_", meta.data = NULL)

  #create seurat mrna object
  expression_matrix_counts_mrna_object<-CreateSeuratObject(t(expression_matrix_counts_mrna), project = "SeuratProject", assay = "RNA", min.cells = 0, min.features = 0, names.field = 1, names.delim = "_", meta.data = NULL)

  #create abseq assay
  RSEC_adt_assay <- CreateAssayObject(counts = t(expression_matrix_counts_abseq), min.cells = 0, min.features = 0)

  #add abseq assay to mrna object
  expression_matrix_counts_mrna_object[["Abseq_RSEC"]] <- RSEC_adt_assay
  expression_matrix_counts_mrna_abseq_object <- expression_matrix_counts_mrna_object

  #add ST information
  smk <- read.table(sprintf("%s/%s/%s", main_dir, dir_in, input_smk_calls), sep = ',', header = TRUE, row.names = 1)
  expression_matrix_counts_mrna_abseq_object <- AddMetaData(expression_matrix_counts_mrna_abseq_object, metadata = smk)

  #remove the multiplets and undetermined from analysis 
  expression_matrix_counts_mrna_abseq_object <- subset(expression_matrix_counts_mrna_abseq_object, subset = Sample_Tag != "Multiplet")
  expression_matrix_counts_mrna_abseq_object <- subset(expression_matrix_counts_mrna_abseq_object, subset = Sample_Tag != "Undetermined")

  #remove PB ST in young donors
  expression_matrix_counts_mrna_abseq_object <- subset(expression_matrix_counts_mrna_abseq_object, subset = Sample_Name != "PB")

  #output object
  return(expression_matrix_counts_mrna_abseq_object)
}
```

#2. Define main directory of input data and cartridge names
```{r define_input_dirs_and_files}
main_dir <- './input_data/csv_data'

#define sub dirs (per age groups)
dir_old <- 'old'
dir_mid_age <- 'mid_age'
dir_young <- 'young'

###data files##

#young
Experiment_1 <- 'Experiment_1_RSEC_MolsPerCell.csv'
Experiment_1_ST <- 'Experiment_1_Sample_Tag_Calls.csv'
Experiment_1_name <- 'Exp_1'


Experiment_2 <- 'Experiment_2_RSEC_MolsPerCell.csv'
Experiment_2_ST <- 'Experiment_2_Sample_Tag_Calls.csv'
Experiment_2_name <- 'Exp_2'


Experiment_3 <- 'Experiment_3_RSEC_MolsPerCell.csv'
Experiment_3_ST <- 'Experiment_3_Sample_Tag_Calls.csv'
Experiment_3_name <- 'Exp_3'


Experiment_4 <- 'Experiment_4_RSEC_MolsPerCell.csv'
Experiment_4_ST <- 'Experiment_4_Sample_Tag_Calls.csv'
Experiment_4_name <- 'Exp_4'


Experiment_5 <- 'Experiment_5_RSEC_MolsPerCell.csv'
Experiment_5_ST <- 'Experiment_5_Sample_Tag_Calls.csv'
Experiment_5_name <- 'Exp_5'

#old age

Experiment_6 <- 'Experiment_6_RSEC_MolsPerCell.csv'
Experiment_6_ST <- 'Experiment_6_Sample_Tag_Calls.csv'
Experiment_6_name <- 'Exp_6'


Experiment_7 <- 'Experiment_7_RSEC_MolsPerCell.csv'
Experiment_7_ST <- 'Experiment_7_Sample_Tag_Calls.csv'
Experiment_7_name <- 'Exp_7'


#mid age

Experiment_8 <- 'Experiment_8_RSEC_MolsPerCell.csv'
Experiment_8_ST <- 'Experiment_8_Sample_Tag_Calls.csv'
Experiment_8_name <- 'Exp_8'


Experiment_9 <- 'Experiment_9_RSEC_MolsPerCell.csv'
Experiment_9_ST <- 'Experiment_9_Sample_Tag_Calls.csv'
Experiment_9_name <- 'Exp_9'

Experiment_10 <- 'Experiment_10_RSEC_MolsPerCell.csv'
Experiment_10_ST <- 'Experiment_10_Sample_Tag_Calls.csv'
Experiment_10_name <- 'Exp_10'


```





#3. Implement Generate Seurat objects for each cartridge
```{r run_anlaysis}


Exp_1_seurat_object <- generate_seurat_object(main_dir, dir_young,
                                            Experiment_1,
                                            Experiment_1_ST)

Exp_2_seurat_object <- generate_seurat_object(main_dir, dir_young,
                                            Experiment_2,
                                            Experiment_2_ST)

Exp_3_seurat_object <- generate_seurat_object(main_dir, dir_young,
                                            Experiment_3,
                                            Experiment_3_ST)

Exp_4_seurat_object <- generate_seurat_object(main_dir, dir_young,
                                            Experiment_4,
                                            Experiment_4_ST)

Exp_5_seurat_object <- generate_seurat_object(main_dir, dir_young,
                                            Experiment_5,
                                            Experiment_5_ST)


Exp_6_seurat_object <- generate_seurat_object(main_dir, dir_old,
                                            Experiment_6,
                                            Experiment_6_ST)

Exp_7_seurat_object <- generate_seurat_object(main_dir, dir_old,
                                            Experiment_7,
                                            Experiment_7_ST)



Exp_8_seurat_object <- generate_seurat_object(main_dir, dir_mid_age,
                                            Experiment_8,
                                            Experiment_8_ST)


Exp_9_seurat_object <- generate_seurat_object(main_dir, dir_mid_age,
                                            Experiment_9,
                                            Experiment_9_ST)

Exp_10_seurat_object <- generate_seurat_object(main_dir, dir_mid_age,
                                            Experiment_10,
                                            Experiment_10_ST)


```

#4. Define main directory of output data
```{r define_output_dir}
#outdir for rds files to store the analysis 
per_cartridge_analysis_rds <- './rds_files/per_cartridge_analysis'
```

#5. Save Seurat objects
```{r save_results}
#save Seurat objects for each cartridge

saveRDS(Exp_1_seurat_object, sprintf("%s/initial_final_%s.rds",per_cartridge_analysis_rds,
                                            Experiment_1_name))

saveRDS(Exp_2_seurat_object, sprintf("%s/initial_final_%s.rds",per_cartridge_analysis_rds,
                                            Experiment_2_name))

saveRDS(Exp_3_seurat_object, sprintf("%s/initial_final_%s.rds",per_cartridge_analysis_rds,
                                            Experiment_3_name))

saveRDS(Exp_4_seurat_object, sprintf("%s/initial_final_%s.rds",per_cartridge_analysis_rds,
                                            Experiment_4_name))

saveRDS(Exp_5_seurat_object, sprintf("%s/initial_final_%s.rds",per_cartridge_analysis_rds,
                                            Experiment_5_name))

saveRDS(Exp_6_seurat_object, sprintf("%s/initial_final_%s.rds", per_cartridge_analysis_rds,             Experiment_6_name))

saveRDS(Exp_7_seurat_object, sprintf("%s/initial_final_%s.rds", per_cartridge_analysis_rds, Experiment_7_name))

saveRDS(Exp_8_seurat_object, sprintf("%s/initial_final_%s.rds", per_cartridge_analysis_rds, Experiment_8_name))

saveRDS(Exp_9_seurat_object, sprintf("%s/initial_final_%s.rds", per_cartridge_analysis_rds, Experiment_9_name))

saveRDS(Exp_10_seurat_object, sprintf("%s/initial_final_%s.rds", per_cartridge_analysis_rds, Experiment_10_name))


```

#Appendix
#Analyze mRNA and AbSeq function - inital settings
```{r}
#mRNA analysis
analyze_mrna <- function(seurat_object){
  DefaultAssay(seurat_object) <- "RNA"
  seurat_object <- NormalizeData(seurat_object)
  seurat_object <- FindVariableFeatures(seurat_object, selection.method = "mvp")
  VariableFeatures(seurat_object)<-rownames(seurat_object[["RNA"]])
  seurat_object <- ScaleData(seurat_object)
  seurat_object <- RunPCA(seurat_object, approx=FALSE, verbose = FALSE)

  nPC_mrna <- 20
  seurat_object <- FindNeighbors(seurat_object, dims = 1:nPC_mrna, verbose = FALSE)
  seurat_object <- FindClusters(seurat_object, resolution = 0.8, verbose = FALSE)
  seurat_object <- RunUMAP(seurat_object, dims = 1:nPC_mrna, verbose = FALSE)
  return(seurat_object)
}

#abseq analysis
analyze_abseq <- function(seurat_object){
  DefaultAssay(seurat_object) <- 'Abseq_RSEC'
  VariableFeatures(seurat_object) <- rownames(seurat_object[["Abseq_RSEC"]])
  seurat_object <- NormalizeData(seurat_object, normalization.method = 'CLR', margin = 2) %>% ScaleData() %>% RunPCA(reduction.name = 'apca', approx = FALSE)
  seurat_object <- RunUMAP(seurat_object, reduction = 'apca', dims = 1:15, assay = 'Abseq_RSEC', reduction.name = 'adt.umap', reduction.key = 'adtUMAP_', verbose = FALSE)
  seurat_object <- RunPCA(seurat_object, npcs = 50, verbose = FALSE, approx = FALSE)
  seurat_object <- FindNeighbors(seurat_object, dims = 1:30, verbose = FALSE)
  seurat_object <- FindClusters(seurat_object, resolution = 0.8, verbose = FALSE)
  return(seurat_object)
}
```

#Analyze mRNA and AbSeq implementation
```{r}
#analyze mRNA modality 
Exp_1_seurat_object <- analyze_mrna(Exp_1_seurat_objectt)
Exp_2_seurat_object <- analyze_mrna(Exp_2_seurat_object)
Exp_3_seurat_object <- analyze_mrna(Exp_3_seurat_object)
Exp_4_seurat_object <- analyze_mrna(Exp_4_seurat_object)
Exp_5_seurat_object <- analyze_mrna(Exp_5_seurat_object)
Exp_6_seurat_object <- analyze_mrna(Exp_6_seurat_object)
Exp_7_seurat_object <- analyze_mrna(Exp_7_seurat_object)
Exp_8_seurat_object <- analyze_mrna(Exp_8_seurat_object)
Exp_9_seurat_object <- analyze_mrna(Exp_9_seurat_object)
Exp_10_seurat_object <- analyze_mrna(Exp_10_seurat_object)


#analyze abseq modality

#NOTE: final seurat_clusters in meta_data will relate to clusters calculated based on abseq params
Exp_1_seurat_object <- analyze_abseq(Exp_1_seurat_object)
Exp_2_seurat_object <- analyze_abseq(Exp_2_seurat_object)
Exp_3_seurat_object <- analyze_abseq(Exp_3_seurat_object)
Exp_4_seurat_object <- analyze_abseq(Exp_4_seurat_object)
Exp_5_seurat_object <- analyze_abseq(Exp_5_seurat_object)
Exp_6_seurat_object <- analyze_abseq(Exp_6_seurat_object)
Exp_7_seurat_object <- analyze_abseq(Exp_7_seurat_object)
Exp_8_seurat_object <- analyze_abseq(Exp_8_seurat_object)
Exp_9_seurat_object <- analyze_abseq(Exp_9_seurat_object)
Exp_10_seurat_object <- analyze_abseq(Exp_10_seurat_object)
```

#Function - assign age groups
```{r}
assign_age_groups <- function(object){
  #create age groups from sample names
  object$sample %>% 
  str_split("1|2|3|4|5|6|8|_") %>% lapply(., "[", 1) %>% as.character() -> age_group_list
  
  #add age group to meta data
  object$age_group <- age_group_list
  
  return(object)
}
```
